---
title: "Module_3.2._Bagging.Rmd"
author: "Katherine Shim"
date: "December 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(caret)
library(ElemStatLearn)
```

###Bootstrap aggregating (bagging)
**Basic idea:**
    1. Resample cases and recalculate predictions
    2. Average or majority vote
**Notes:**
    1. Similar bias
    2. Reduced variance
    3. More useful for non-linear functions
    
###Example: Ozone data
Each ith row of ll matrix is the prediction from the loess curve based on ith resampled data.
```{r}
data(ozne, package = "ElemStatLearn")
ozone <- ozone[order(ozone$ozone),]
ll <- matrix(NA, nrow=10, ncol=155)
for (i in 1:10){
  ss <- sample(1:dim(ozone)[1], replace=T)
  ozone0 <- ozone[ss,]
  ozone0 <- ozone0[order(ozone0$ozone),]
  #loess function fit a smooth curve with given data
  loess0 <- loess(temperature ~ ozone, data=ozone0, span=0.2)
  ll[i,] <- predict(loess0, newdata=data.frame(ozone=1:155))
  }
```
  
###Bagged loess
Then I will average the 10 predictions from 10 different loess curve based on 10 resampled dataset.
```{r}
#Plot the sample data
plot(ozone$ozone, ozone$temperature, pch=19, cex=0.5)
#Plot 10 loess curves
for (i in 1:10) { lines(1:155, ll[i,], col="grey", lwd=2)}
#Average the loess curve. apply funciton has 2 to average of each column over rows
lines(1:155, apply(11,2,mean), col="red", lwd=2)
```

###Bagging in caret

* Some models perform baggin for you , in `train` function consider `method` options
    + `bagEarth`
    + `treebag`
    + `bagFDA`
* Aternatively you can bag any model you choose using the `bag` function

###More bagging in caret
You can build your own bag model
```{r}
predictors <- data.frame(ozone=ozone$ozone)
temperature = ozone$temperature
treebag <- bag(predictors, temperature, B = 10,
bagControl = bagControl(fit = ctreeBag$fit, predict = ctreeBag$pred, aggregate = ctreeBag$aggregate))
plot(ozone$ozone, temperature, col='lightgrey', pch=19)
points(ozone$ozone, predict(treebag$fit[[1]]$fit, predictors), pch=19, col="red")
points(ozone$ozone, predict(treebag, predictors), pch=19, col="blue")
```
