---
title: "R Classes and Methods"
author: "Katherine Shim"
date: "November 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Classes and Methods

* A system for doing object oriented programming
* R has two styles of classes and methods
    1. S3 classes/methods (old school)
    2. S4 classes/methods: The code for implementing S4 classes/methods in R is in the methods package, which is usually loaded by default

### Object Oreinted Programming in R

* A class is a desciprtion of an thing. A class can be defined using `setClass()` in the methods package
* An object is an instance of a class. Objects can be created using `new()`
* A method is a function that only operates on a certain class of obejcts.
* A generic function is an R function which dispatches methods. A generic function typically encasulates a "generic" concept (e.g. `plot`, `mean`, `predict`,...)
* A method is the implementation of a generic function for an obejct of a particular class.


### Classes
All objects in R have a class which can be determined by the class function
```{r}
class(1)
class(TRUE)
class(rnorm(100))
class(NA)
class("foo")
x <- rnorm(100)
y <- x + rnorm(100)
fit <- lm(y~x)
class(fit)
```

### Generics/Methods in R

* S4 and S3 style generic functions look different but conceptually, they are the same.
* You can write new methods for an existic generic OR create your own generics and associated methods.
* If a data type doesn't exist in R that matches programmers needs, the programmer can define a new class along with generics/methods that go with it.

### S3 generic and methods in R
```{r}
#These generic functions call UseMethod("function")
mean
print
#The mean generic function has a number of methods associated with it
methods("mean")
```

### S4 generic and methods in R
The `show` function is from the methods package and is the S4 equivalent of `print`
```{r}
show
showMethods("show")
```

###Generic/method mechanism
The first argument of a generic function is an object of a particular class

1. The generic function checks the class of the object
2. A search is done to see if there is an appropriate method for that class.
3. If there exists a method for that class, then that method is called on the object and we're done.
4. If a method for that class doesn't exist, a search is done to see if there is a default method for the generic. If a default exists, then the default method is called.
5. If a default method doesn't exist, then an error is thrown.

### S3 Example1

1. The class of x is numeric
2. There's no mean method for numeric objects
3. It calls the default function for `mean`
```{r}
set.seed(2)
x <- rnorm(100)
mean(x)
```
