---
title: "Leaflet"
author: "Katherine Shim"
date: "October 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```

Leaflet is a Javascript libraries fo creating interactive maps. It allows you to create your own leaflet maps without knowing any Javascript.

###First basic example
```{r}
#The leaflet() function creates a map widget that you can store in a variable
my_map <- leaflet() %>% 
  #addTiles() function adds mapping data from Open street map
  addTiles()
my_map
```

###Add markers
```{r, eval=FALSE}
my_map <- my_map %>%
  addMarkers(lat=31, lng=-52, 
             popup="Middle of nowhere")
my_map
```

###Adding multiple markers
```{r}
set.seed(2016-04-25)
df <- data.frame(lat=runif(20, min=39.2, max=39.3),
                 lng = runif(20, min=-76.6, max=-76.5))
df %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers()

```

###Adding custom markers
```{r, eval=FALSE}
hopkinsIcon <- makeIcon(
  iconUrl = "https://cdn2.iconfinder.com/data/icons/holloween/48/Halloween-09-512.png",
  iconWidth = 31*215/230, iconHeight = 31,
  iconAnchorX = 31*215/230/2, iconAnchorY = 16
)
hopkinsLatLong <- data.frame(
  lat = c(39.2973166, 39.3288851, 39.2906617),
  lng = c(-76.5929798, -76.6206598, -76.5469683))
hopkinsSites <- c(
  "<a href='http://www.jhsph.edu/'>East Baltimore Campus</a>",
  "<a href='https://apply.jhu.edu/visit/homewood/'>Homewood Campus</a>",
  "<a href='http://www.hopkinsmedicine.org/johns_hopkins_bayview/'>Bayview Medical Center</a>",
  "<a href='http://www.peabody.jhu.edu/'>Peabody Institute</a>",
  "<a href='http://carey.jhu.edu/'>Carey Business School</a>"
)
hopkinsLatLong %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon = hopkinsIcon, popup = hopkinsSites)
```

###Cluster markers
When you zoom into each cluster, the clusters will separate until you can see the individual markers.
```{r}
df <- data.frame(lat = runif(500, min = 39.25, max = 39.35),
                 lng = runif(500, min = -76.65, max = -76.55))
df %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(clusterOptions = markerClusterOptions())
```

### Circle markers
```{r}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55))
df %>% 
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers()
```

###Drawing circles
You can draw arbitrary shapes on the maps you create. The code below draws a map where the circle on each city is proportional to the population of that city.
```{r, eval=FALSE}
md_cities <- data.frame(name = c("Baltimore", "Frederick", "Rockville", "Gaithersburg", 
                                 "Bowie", "Hagerstown", "Annapolis", "College Park", "Salisbury", "Laurel"),
                        pop = c(619493, 66169, 62334, 61045, 55232,
                                39890, 38880, 30587, 30484, 25346),
                        lat = c(39.2920592, 39.4143921, 39.0840, 39.1434, 39.0068, 39.6418, 38.9784, 38.9897, 38.3607, 39.0993),
                        lng = c(-76.6077852, -77.4204875, -77.1528, -77.2014, -76.7791, -77.7200, -76.4922, -76.9378, -75.5994, -76.8483))
md_cities %>%
  leaflet() %>%
  addTiles() %>%
  addCircles(weight = 1, radius = sqrt(md_cities$pop) * 30)
```

###Drawing rectangles
```{r}
leaflet() %>%
  addTiles() %>%
  addRectangles(lat1 = 37.3858, lng1 = -122.0595, 
                lat2 = 37.3890, lng2 = -122.0625)
```

###Adding legends
```{r, message=FALSE}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55),
                 col = sample(c("red", "blue", "green"), 20, replace = TRUE),
                 stringsAsFactors = FALSE)
df %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(color = df$col) %>%
  addLegend(labels = LETTERS[1:3], colors = c("blue", "red", "green"))
```