---
title: "Hypothesis Testing"
output: html_document
header-includes:
    - \usepackage{setspace}\doublespacing
---

```{r setup, include=FALSE}
library(UsingR)
library(datasets); data(ChickWeight); library(reshape2)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

##Hypothesis testing

* Hypothesis testing is concerned with making decisions using data.
* A null hypothesis is specified that represents the status quo, usually labeled $H_0$
* The null hypothesis is assumed true and statistical evidence is required to reject it in favor of an alternative hypothesis.
* The alternative hypotheses are typically of the form <, > or not !=.
* Note that there are four possible outcomes of our statistical decision process.
    1. Correctly accept null.
    2. Falsely reject null hypothesis (Type I error).
    3. Correctly reject null.
    4. Falsely accept null (Type II error).
* A reasonable strategy would reject the null hypothesis if $\overline{X}$ was larger than some constant C.
* Typically, C is chosen so that the probability of a Type I error, $\alpha$ is 0.05.
* $\alpha$ = Type I error rate = Probability of falsely rejecting the null hypothesis.
  
**Example**
A respiratory disturbance index of more than 30 events/hour is considered evidence of severe sleep disordered breathing (SDB). Suppose that in a sample of 100 overweight subjects with other risk factors for SDB at a sleep clnic, the mean RDI was 32 events/hour with a standard deviation of 10 events/hour.  

* We might want to test the hypothesis that
    + $H_0 : \mu = 30$
    + $H_a : \mu >30$
    + where $\mu$ is the population mean RDI.
* Standard error of the mean is $1/\sqrt{100}=1$
* Under $H_0 \overline{X} \sim N(30, 1)$, we want to choose C so that the $P(\overline{X}>C; H_0)$ is 5%.
* The 95th percentile of z-score is 1.645. So, C=30+1*1.645 = 31.645
    + Then the probability that a N(30, 1) is larger than it is 5%
    + So the rule "Reject $H_0$ when $\overline{X} >= 31.645$" has the property that the probability of rejection is 5% when $H_0$ is true (for the $\mu_0, \sigma$ and $n$ given)
* We would reject because the Z-score from the sample is \[\frac{32-30}{10/\sqrt(100)}=2\]
which is greater than 1.645. 
  
###T tests

* Consider the example above except n= 16 instead of 100. 
* The statistic
\[\frac{\overline{X}-30}{s/\sqrt{16}}\]
follows a T distribution with 15 df under $H_0$. So our test statistic is now $\sqrt{16}(32-30)/10=0.8$
* The 95th percentile of the T distribution with 15 df is 1.7531 (obtained via qt(0.95, 15)) 
* Since 0.8 < 1.7531, we fail to reject.

**Two sided tests**  

* Consider the same example. Suppose that we would reject the null hypothesis if the mean is too large or too small.
* $H_a : \mu \neq 30$
* Then we want the probability of rejecting under the null to be 5%, split equally as 2.5% in the upper and lower tail.
* Thus, we will reject the null hypothesis if test statistic is larger than 2.1314 from qt(0.975, 15).
* Since test statistic is 0.8, we fail again. But it is obvious to fail two sided test because we already failed the one sided test.

**T test in R**
```{R}
data(father.son)
t.test(father.son$sheight-father.son$fheight)
```
  
**Connections with confidence intervals**

* Consider testing $H_0 : \mu = \mu_0$ versus $H_a : \mu \neq \mu_0$
* Take the set of all possible values for which you fail to reject $H_0,$ this set is a $(1-\alpha)*100%$ confidence interval fro $\mu$
* The same works in reverse: if a $(1-\alpha)*100%$ interval contains $\mu_0$, then we fail to reject $H_0$
  
**Two group testing**
* Test $H_0 \mu_1 = \mu_2$
```{r}
##define weight gain or loss
wideCW <- dcast(ChickWeight, Diet + Chick ~ Time, value.var = "weight")
names(wideCW)[-(1 : 2)] <- paste("time", names(wideCW)[-(1 : 2)], sep = "")
wideCW <- mutate(wideCW,
  gain = time21 - time0
)
wideCW14 <- subset(wideCW, Diet %in% c(1, 4))
t.test(gain ~ Diet, paired = FALSE, 
       var.equal = TRUE, data = wideCW14)
```

###Exact binomial test
* Suppose a friend has 8 children, 7 of which are girls and none are twins. Perform the relevant hypothesis test. $H_0 : p =0.5 H_a: p>0.5$ What's the relevant rejection region so that the probability of rejecting is (less than) 5%?  
When there's 7 or more girls, then the type 1 error rate is smaller than 5%.  
**Notes**

1. It's impossible to get an exact 5% level test for this case due to the discreteness of the binomial (you can't have fraction of child).
    1.1. The closest is the rejection region [7:8]
    1.2. Any alpha level lower than 0.0039 is not attainable.
2. For larger sample sizes, we could do a normal approximation.
3. Two sided test isn't obvious.
    3.1. Given a way to do two sided tests, we could take the set of values of $p_0$ for which we fail to reject to get an exact binomial confidence interval (called the Clopper/Pearson).
4. For these problems, people always create a P-value rather than computing the rejection region.
