---
title: "Variability"
output: html_document
header-includes:
    - \usepackage{setspace}\doublespacing
---

```{r setup, include=FALSE,echo=FALSE}
library(UsingR)
```

##Variance
The variance of a random variable is a measure of spread. The square root of the variance is standard deviation.
\[Var(X) = E[(X-\mu)^2] = E[X^2]-E[X]^2\]
**Sample variance**
\[S^2 = \frac{\sum_{i-1}(X_i-\overline{X})^2}{n-1}\]
The expected value of the sample variance is estimated of the population variance. This estimation gets better as n gets larger. Sample variance is also unbiased estimation of population variance.

###Standard error
We call the standard deviation of a statistic a standard error.
\[Var(\overline{X}) = \sigma^2/n\]
In this scenario, the standard deviation of the sample mean is standard error of the mean. Likewise standard error of a regression coefficient talks about the variabilities of the regression coefficient. Since sample variance is good estimator for population variance, a logical estimate of standard error of a mean is $s/\sqrt{n}$.  

* $s/\sqrt{n}$, the standard error, talks about how variable averages of random samples of size n from the population are.

###Example
```{r}
data(father.son)
x <- father.son$sheight
n <- length(x)
#Estimate for population son's height. If we assume that x is the random sample from the population.
#var(x) and sd(x) are estimates of variability of population
#var(x)/n and sd(x)/sqrt(n) are estimates of the variability in averages of son's height.
round(c(var(x), var(x)/n, sd(x), sd(x)/sqrt(n)),2)
```
