---
title: "Distributions"
output: html_document
header-includes:
- \doublespacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Bernoulli distributions
The Bernoulli distribution arises as the result of a binary outcome. Bernoulli random variables take only the values 1 and 0 with probabilities of $p$ and $1-p$ respectively.  
**pmf**: $P(X=x)=p^x(1-p)^{1-x}$  
**mean**: $p$  
**variance**: $p(1-p)$  
  
###Binomial distributions
Let $X_1,..,X_n$ be iid Bernoulli(p), then $X=\sum_{i=1}{n}X_i$ is a binomial random variable.  
**pmf**: $P(X=x)= nCx* p^x(1-p)^{n-x}$  
**example**: Suppose a friend has 8 children, 7 of which are girls. If each gender has an independent 50% probability for each birth, what's the probability of getting 7 or more girls out of 8 births?  
=> $8C7*0.5^7(1-0.5)^1+8C8*0.5^8(1-0.5)^0 = 0.04$
```{r}
pbinom(6, size=8, prob=0.5, lower.tail=FALSE)
```
****
  
###Normal distribution
Gaussian distribution with mean $\mu$ and variance $\sigma^2$. $E[x]=\mu$ and $Var(X)=\sigma^2$, $X \sim N(\mu, \sigma^2)$. When $\mu=0$ and $\sigma=1$, the resulting distribution is standard normal distribution.  
**Facts about the normal density**  
* If $X \sim N(\mu, \sigma^2)$ then
\[Z=\frac{X-\mu}{\sigma} \sim N(0,1)\]  
* If Z is standard normal
\[X=\mu+\sigma Z \sim N(\mu, \sigma^2)\]
**Example1**: What is the 95th percentile of a $N(\mu, \sigma^2)$ distributions?
```{r}
mu=0
sd=1
qnorm(0.95, mu, sd)
```
**Example2**: Assume that the number of daily ad clicks for a company is (approximately) normally distributed with a mean of 1020 and a standard deviation of 50. What's the probability of getting more than 1,160 clicks in a day?
```{r}
mu=1020
sd=50
pnorm(1160, mu, sd, lower.tail=FALSE)
```
**Example3**: Assume that the number of daily ad clicks for a company is (approximately) normally distributed with a mean of 1020 and a standard deviation of 50. What number of daily ad clicks would represent the one where 75% of days have fewer clicks (assuming days are independent and identically distributed?)
```{r}
mu=1020
sd=50
qnorm(0.75, mu, sd)
```
  
###Poisson distribution
**Usage**  
1. Modeling count data  
2. Modeling event-time or survival data
3. Modeling contingency tables
4. Approximating binomials when n is large and p is small

**pmf**: $P(X=x;\lambda)=\frac{\lambda^x e^{-\lambda}}{x!}$  
**mean**: $\lambda$  
**variance**: $\lambda$  

**Rates and Poisson random variables**  
$X \sim Poisson(\lambda t)$ where

* $\lambda = E[X/t]$ is the expected count per unit of time
* $t$ is the total monitoring time  
*Example: The number of people that show up at a bus stop is Poisson with a mean of 2.5 per hour. If watching the bus stop for 4 hours, what is the probability that 3 or fewer people show up for the whole time?
```{r}
ppois(3, lambda=2.5*4)
```
  
**Poisson approximation to the binomial**  

* When n is large and p is small the poisson distribution is an accurate approximation to the binomial distribution
* Notation
- $X \sim Binomial(n,p)$
- $\lambda = np$
- n gets large
- p gets small
* Example: We flip a coin with success probability 0.01 five hundred times. What's the probability of 2 or fewer successes?
```{r}
pbinom(2, 500, 0.01)
ppois(2, lambda=500*0.01)
```
