---
title: "Pitch Presentation"
author: "Katherine"
date: "November 6, 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Topic

- This R markdown presentation provides reproducible pitch slides that explains the Shiny application I created.
- In order to find the Shiny app, please go to https://salus0324.shinyapps.io/Courseproject/.

## Summary of the Shiny Application

- Purpose: Perform univariate linear regression analysis on the Mtcars dataset to predict mpg using the chose predictor variable.
- Predictor variable choices:
    + Cylinders (cyl)
    + Displacement (disp)
    + Gross horsepower (hp)
    + Weight (wt)
    + Carburetors (carb)
    + Gears (gear)
- Output:
    + Scatter plot of the data
    + Linear regression model
    + LM slope and intercept

## Example 1
If user chose Cylinders (cyl) as the predictor variable the output plot would look like below.

```{r cars, echo = FALSE}
fit <- lm(mtcars$mpg ~ mtcars$cyl, data = mtcars)
plot(mtcars$cyl, mtcars$mpg, xlab = "cyl", ylab = "MPG", main = "Linear model", cex = 1.5, pch = 16, bty = "n")
abline(fit, col = "blue", lwd = 2)
```

## Example 2
If user chose Displacement (disp) as the predictor variable the output plot would look like below.
```{r pressure}
fit <- lm(mtcars$mpg ~ mtcars$disp, data = mtcars)
plot(mtcars$disp, mtcars$mpg, xlab = "disp", ylab = "MPG", main = "Linear model", cex = 1.5, pch = 16, bty = "n")
abline(fit, col = "blue", lwd = 2)
```

