---
title: "Forecasting"
author: "Katherine Shim"
date: "December 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(quantmod)
library(forecast)
```

### What is different?

* Data are dependent over time
* Specific pattern types
  * Trends - long term increase or decrease
  * Seasonal patterns - patterns related to time of week, month, year, etc.
  * Cycles - patterns that rise and fall periodically
* Subsampling into training/test is more complicated
* Similar issues arise in spatial data 
  * Dependency between nearby observations
  * Location specific effects
* Typically goal is to predict one or more observations into the future. 
* All standard predictions can be used (with caution!)

### Example - Google stock price
```{r}
from.dat <- as.Date("01/01/08", format="%m/%d/%y")
to.dat <- as.Date("12/31/13", format="%m/%d/%y")
getSymbols("GOOG", src="yahoo", from = from.dat, to = to.dat)
head(GOOG)
```

### Summarize monthly and store as time series

```{r}
mGoog <- to.monthly(GOOG)
# Get opening price
googOpen <- Op(mGoog)
# Create a time series object
ts1 <- ts(googOpen,frequency=12)
plot(ts1,xlab="Years+1", ylab="GOOG")
```

### Decompose a time series into parts

```{r}
plot(decompose(ts1),xlab="Years+1")
```

###Training and test sets
```{r}
ts1Train <- window(ts1,start=1,end=5)
ts1Test <- window(ts1,start=5,end=(7-0.01))
ts1Train
```

### Simple moving average
\[ Y_{t}=\frac{1}{2*k+1}\sum_{j=-k}^k {y_{t+j}}\]

```{rs}
plot(ts1Train)
lines(ma(ts1Train,order=3),col="red")
```

### Exponential smoothing
\[\hat{y}_{t+1} = \alpha y_t + (1-\alpha)\hat{y}_{t-1}\]
The data points near to current time has heavier weight than the ones further back.
```{r}
ets1 <- ets(ts1Train,model="MMM")
fcast <- forecast(ets1)
# The shade on the plot is the prediction bound
plot(fcast); lines(ts1Test,col="red")
#Get the accuracy
accuracy(fcast, ts1Test)
```

