---
title: "Intro to regression and least squares"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(UsingR)
library(reshape)
library(dplyr)
```

## Galton's data - child and parents height
```{r}
data(galton)
long <- melt(galton)
g <- ggplot(long, aes(x=value, fill=variable))
g <- g + geom_histogram(colour = "black", binwidth=1)
g <- g + facet_grid(.~variable)
g
```
  
###Least squares - Children's height

* Consider only the children's height.
    + The middle is the value $\mu$ that minimizes \[\sum_{i=1}^{n} (Y_i-\mu)^2\]
* This is al so physical center of mass of the histogram.
* $\mu=\overline{Y}$
```{r}
g <- ggplot(galton, aes(x=child))
g <- g + geom_histogram(fill="salmon", colour="black", binwidth=1)
g <- g + geom_vline(xintercept = mean(galton$child), size=3)
g
```
  
###Exploration of Children and parents heights
Scatterplot
```{r}
freqData <- as.data.frame(table(galton$child, galton$parent))
names(freqData) <- c("child", "parent", "freq")
freqData$child <- as.numeric(as.character(freqData$child))
freqData$parent <- as.numeric(as.character(freqData$parent))
g <- ggplot(filter(freqData, freq > 0), aes(x = parent, y = child))
g <- g  + scale_size(range = c(2, 20), guide = "none" )
g <- g + geom_point(colour="grey50", aes(size = freq+20, show_guide = FALSE))
g <- g + geom_point(aes(colour=freq, size = freq))
g <- g + scale_colour_gradient(low = "lightblue", high="white")                    
g
```
  
###Regression through the origin

* Suppose that $X_i$ are the parents' heights
* Consider picking the slope $\beta$ that minimizes \[\sum_{i=1}^{n}(Y_i-X_i\beta)^2\]
* This is using the origin as a pivot point picking the line that minimizes the sum of the squared vertical distances of the points to the line.
* Subtract the means so that the origin is the mean of the parents and children's heights
```{r}
# -1 is to get rid of the intercept
lm(I(child-mean(child))~I(parent-mean(parent))-1, data=galton)
```
We subtracted the mean to center the data to origin. The slope $\beta$ is 0.6463.