---
title: "Probability"
output: html_document
header-includes:
    - \usepackage{setspace}\doublespacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Probability
Given a random experiment a probability measure is a population quantity that summarizes the randomness.  
  
###Rules of probability:

* The probability that nothing occurs is 0
* The probability that something occurs is 1
* The probability of something is 1 minus the probability that the opposite occurs
* The probability of at least one of two things that can not simultaneously occur (mutually exclusive) is the sum of their respective probabilities
* If an event A implies the occurrence of event B, then the probability of A occurring is less than the probability that B occurs
* For any two events the probability that at least one occurs is the sum of their probabilities minus their intersection  
  
###Probability mass/density function (pmf/pdf)  
**Random variable**: a discrete or continuous numerical outcome of an experiment.  
**Probability mass function (pmf)**: A probability mass function evaluated at a value corresponds to the probability that a random variable takes that value. To be a valid pmf a function, p, must satisfy:  

* It must always be larger than or equal to 0.
* The sum of the possible values that the random variable can take has to add up to one.  

**Probability density function (pdf)**: A probability density function is a function associated with a continuous random variable. Areas under pdfs correspond to probabilities for that random variable. To be a valid pdf a function, p, must satisfy:  

* It must be larger than or equal to zero everywhere.
* The total area under it must be one.  

**Cumulative distribution function (cdf)**: The cumulative distribution function of a random variable, X, returns the probability that the random variable is less than or equal to the value x. \[F(x) = P(X <= x)\]

**Survival function**: The survival function of a random variable X is defined as the probability that the random variable is greater than the value x. \[S(x) = P(X > x) = 1 - F(x)\]

**Quatiles**: The $\alpha$ quantile of a distribution with distribution function F is the point $x_\alpha$ so that $F(x_\alpha) = \alpha$.

* A percentile is simply a quantile with $\alpha$ expressed as a percent
* The median is the 50th percentile

###R functions
```{r }
#Find the percentage under the given point
pbeta(0.7071068, 2, 1) #50% under 0.707
#Find the point with given quantile.
qbeta(0.5,2,1) #Median is 50th quantile
```
